---
output:
  pdf_document:
    highlight: tango
    toc: yes
  html_document:
    keep_md: yes
---

```{r config, echo=FALSE}
library(knitr)
opts_chunk$set(background='gray80', tidy=FALSE, cache=FALSE, keep_md=TRUE, 
               fig.path='figure/', fig.height=7, fig.width=7)
```


# BEMA Figures for comparing distance metrics using HTS bacteria. 



```{r loadPackages}
library(wesanderson)
library(scales)
library(vegan)
```







```{r inputData}
# setwd('~/Dropbox/BEMA/bacteriaHTS/closed_ref_pickedOTUs/BEMAFinalAnalysis/R/ordinations')
ordBCIn <- read.delim(pipe('cut -f1-4 ordinationsFromQIIME/beta100_bc/bray_curtis_pc.txt'), 
                    head=TRUE, row.names=1)
ordUNIIn <- read.delim(pipe('cut -f1-4 ordinationsFromQIIME/beta100_ununif/unweighted_unifrac_pc.txt'), 
                     head=TRUE, row.names=1)
taxaBCIn <- read.delim(pipe('cut -f1-4 ordinationsFromQIIME/beta100_bc/3dplots_bc_L6_points.txt'), 
                     head=TRUE, row.names=1)
taxaUNIIn <- read.delim(pipe('cut -f1-4 ordinationsFromQIIME/beta100_ununif/3dplots_ununif_L6_points.txt'), 
                      head=TRUE, row.names=1)

bokClosedIn <- read.delim(pipe('cut -f1-4 openClosed/bokulich_ununi_closed_pc.txt'), 
                        head=TRUE, row.names=1)
bokOpenIn <- read.delim(pipe('cut -f1-4 openClosed/bokulich_ununi_open_pc.txt'), 
                        head=TRUE, row.names=1)
lfClosedIn <- read.delim(pipe('cut -f1-4 openClosed/leff_flores_ununi_closed_pc.txt'), 
                        head=TRUE, row.names=1)
lfOpenIn <- read.delim(pipe('cut -f1-4 openClosed/leff_flores_ununi_open_pc.txt'), 
                        head=TRUE, row.names=1)
meaClosedIn <- read.delim(pipe('cut -f1-4 openClosed/meadow_ununi_closed_pc.txt'), 
                        head=TRUE, row.names=1)
meaOpenIn <- read.delim(pipe('cut -f1-4 openClosed/meadow_ununi_open_pc.txt'), 
                        head=TRUE, row.names=1)


bokOpenDivIn <- read.delim('openClosed/bokOpenAlphaDiv.txt', row.names=1)
bokClosedDivIn <- read.delim('openClosed/bokClosedAlphaDiv.txt', row.names=1)

load('../countData/dfWithColors.RData')
```


Remove extra info from PC dataframe. This has `eigvals` and `% variation explained` as the last two rows. Create a function to make new cleaned datasets. 

```{r cutEigs}
cutEigs <- function(ord, ordName = 'newOrd', eigName = 'eigs') {
  assign(eigName, ord[c(nrow(ord)-1, nrow(ord)), ], pos=1)
  assign(ordName, ord[c(1:(nrow(ord)-2)), ], pos=1)
}
# cutEigs(bokOpenIn)
```


Big dataset - Bray Curtis vs UniFrac. 

```{r}
cutEigs(ord=ordBCIn, ordName='ordBC', eigName='eigBC')
cutEigs(ord=ordUNIIn, ordName='ordUNI', eigName='eigUNI')
```



Bokulich Dairy and NICU, closed and open reference datasets. 

```{r cutEigsOpenClosed}
cutEigs(ord=bokClosedIn, ordName='bokClosed', eigName='eigBokClosed')
cutEigs(ord=bokOpenIn, ordName='bokOpen', eigName='eigBokOpen')
```




Leff and Flores Kitchens, closed and open reference datasets. 

```{r cutLFEigsOpenClosed}
cutEigs(ord=lfClosedIn, ordName='lfClosed', eigName='eigLFClosed')
cutEigs(ord=lfOpenIn, ordName='lfOpen', eigName='eigLFOpen')
```



Meadow air and surfaces, closed and open reference datasets. 

```{r cutMeadowEigsOpenClosed}
cutEigs(ord=meaClosedIn, ordName='meaClosed', eigName='eigMeaClosed')
cutEigs(ord=meaOpenIn, ordName='meaOpen', eigName='eigMeaOpen')
```




All samples are present but in a different order. 




```{r checkAlignOpenClosed}
dim(bokClosed); dim(bokOpen)
all(row.names(bokOpen) %in% row.names(bokClosed))
bokOpen <- bokOpen[row.names(bokClosed), ]
identical(row.names(bokOpen), row.names(bokClosed))
```


```{r checkLFAlignOpenClosed}
dim(lfClosed); dim(lfOpen)
all(row.names(lfOpen) %in% row.names(lfClosed))
all(row.names(lfClosed) %in% row.names(lfOpen))
setdiff(row.names(lfOpen), row.names(lfClosed))  # one extra sample
lfOpen <- lfOpen[row.names(lfClosed), ]
identical(row.names(lfOpen), row.names(lfClosed))
```


```{r checkMeadowAlignOpenClosed}
dim(meaClosed); dim(meaOpen)
all(row.names(meaClosed) %in% row.names(meaOpen))  # one extra sample
meaOpen <- meaOpen[row.names(meaClosed), ]
identical(row.names(meaOpen), row.names(meaClosed))
```





```{r makeOCMaps}
all(row.names(bokOpen) %in% row.names(dat))
bokMap <- dat[row.names(bokOpen), ]
lfMap <- dat[row.names(lfOpen), ]
meaMap <- dat[row.names(meaOpen), ]
```







```{r procrustes}
bokCO <- procrustes(bokClosed, bokOpen)
bokOC <- procrustes(bokOpen, bokClosed)
bokTest <- protest(bokOpen, bokClosed)
# 
# plot(bokOC, kind=1)
# 
# plot(bokOC, kind=0, las=1)
# lines(bokOC, col='gray20', type='arrows', code=1, length=.08)
# points(bokOC$Yrot, pch=21, bg=bokMap$col, col='gray20', lwd=.3)  # From points
```


```{r procrustesLF}
lfCO <- procrustes(lfClosed, lfOpen)
lfOC <- procrustes(lfOpen, lfClosed)
lfTest <- protest(lfOpen, lfClosed)
# 
# plot(lfOC, kind=1)
# 
# plot(lfOC, kind=0, las=1)
# lines(lfOC, col='gray20', type='arrows', code=1, length=.08)
# points(lfOC$Yrot, pch=21, bg=lfMap$col, col='gray20', lwd=.3)  # From points
```



```{r procrustesMeadow}
meaCO <- procrustes(meaClosed, meaOpen)
meaOC <- procrustes(meaOpen, meaClosed)
meaTest <- protest(meaOpen, meaClosed)

# plot(meaOC, kind=1)
# 
# plot(meaOC, kind=0, las=1)
# lines(meaOC, col='gray20', type='arrows', code=1, length=.08)
# points(meaOC$Yrot, pch=21, bg=meaMap$col, col='gray20', lwd=.3)  # From points
```


Trying to find out why one study was much different closed vs open. Nothing convincing here. 

```{r}
dim(bokOpenDivIn)
dim(bokClosedDivIn)
all(row.names(bokMap) %in% row.names(bokClosedDivIn))
bokMap$closed <- bokClosedDivIn[row.names(bokMap), 1]
bokMap$open <- bokOpenDivIn[row.names(bokMap), 1]
# plot(bokMap$closed, bokMap$open, pch=21, bg=bokMap$col)
# segments(0,0,1000,1000)
# boxplot(bokMap$open/bokMap$closed ~ factor(bokMap$phinchID))
```






bokMap[, c('phinchID', 'col')]
points(.2, .15, pch=21, bg='#FD6467')

```{r procrustes3panel}

pdf('procrustes3panel.pdf', width=4, height=10, useDingbats=FALSE)
par(mfrow=c(3,1))
plot(bokOC, kind=0, las=1, main='(a) CA Dairy and NICU')
lines(bokOC, type='arrows', code=1, length=.03)
points(bokOC$Yrot, pch=21, bg=bokMap$col, lwd=.3)  # From points
legend('bottomright', legend=c('CA NICU', 'CA Dairy'), 
       pch=21, pt.bg=unique(bokMap$col), bty='n', pt.cex=2)
mtext(paste('p =', bokTest$signif, '  '), 
      side=3, adj=1, line=-1.5, cex=.7)
mtext(paste('corr =', format(bokTest$t0, digits=2), '  '), 
      side=3, adj=1, line=-2.5, cex=.7)
# unique(meaMap$phinchID)
# unique(meaMap$col)

plot(lfOC, kind=0, las=1, main='(b) NC Homes and CO Residential Kitchens')
lines(lfOC, type='arrows', code=1, length=.03)
points(lfOC$Yrot, pch=21, bg=lfMap$col, lwd=.3)  # From points
legend('topleft', legend=c('NC Homes', 'CO Residential Kitches'), 
       pch=21, pt.bg=unique(lfMap$col), bty='n', pt.cex=2)
mtext(paste('p =', lfTest$signif, '  '), 
      side=3, adj=1, line=-1.5, cex=.7)
mtext(paste('corr =', format(lfTest$t0, digits=2), '  '), 
      side=3, adj=1, line=-2.5, cex=.7)



plot(meaOC, kind=0, las=1, main='(c) OR Classroom Air and Surfaces')
lines(meaOC, type='arrows', code=1, length=.03)
points(meaOC$Yrot, pch=21, bg=meaMap$col, lwd=.3)  # From points
legend('bottomleft', legend=c('OR Classroom Air', 'OR Classroom Surfaces'), 
       pch=21, pt.bg=unique(meaMap$col), bty='n', pt.cex=2)
mtext(paste('p =', meaTest$signif, '  '), 
      side=1, adj=1, line=-2.5, cex=.7)
mtext(paste('corr =', format(meaTest$t0, digits=2), '  '), 
      side=1, adj=1, line=-1.5, cex=.7)



dev.off()
```













```{r checkMap}
head(dat)  # Description
head(ordBC)
dim(dat)
dim(ordBC)
dim(ordUNI)
all(row.names(ordBC) %in% row.names(ordUNI))
all(row.names(ordBC) %in% row.names(dat))
setdiff(row.names(ordBC), row.names(dat))
```



```{r}
sameNames <- intersect(row.names(ordBC), row.names(dat))
ordBC <- ordBC[sameNames, ]
ordUNI <- ordUNI[sameNames, ]
dat <- dat[sameNames, ]
identical(row.names(ordBC), row.names(dat))
```



```{r fixTaxonNamesBC}
taxaBC <- taxaBCIn
namesBC.tmp <- row.names(taxaBCIn)
namesBC.tmp <- gsub('\\;[a-z]\\_\\_$', '', namesBC.tmp)  # run twice to kill nonames
namesBC.tmp <- gsub('\\;[a-z]\\_\\_$', '', namesBC.tmp)
namesBC <- gsub('(.+)\\;[a-z]\\_\\_([[:alpha:]]{1,})$', '\\2', namesBC.tmp)
namesBC <- gsub('Enterobacteriaceae', 'Enterobact.', namesBC)
```


```{r fixTaxonNamesUNI}
taxaUNI <- taxaUNIIn
namesUNI.tmp <- row.names(taxaUNI)
namesUNI.tmp <- gsub('\\;[a-z]\\_\\_$', '', namesUNI.tmp)  # run twice to kill nonames
namesUNI.tmp <- gsub('\\;[a-z]\\_\\_$', '', namesUNI.tmp)
namesUNI <- gsub('(.+)\\;[a-z]\\_\\_([[:alpha:]]{1,})$', '\\2', namesUNI.tmp)
namesUNI <- gsub('Enterobacteriaceae', 'Enterobact.', namesUNI)
```







```{r offsetNames}
posBC <- c(4, 4, 2, 2, 2, 4, 4, 2, 2, 4)
posUNI <- c(4, 4, 4, 2, 4, 2, 2, 2, 2, 2)
```






```{r moreColors}
dat$colBuildingType <- ''
dat$colMatrix <- ''

dat$Building_Type3 <- dat$Building_Type
dat$Building_Type3[dat$Building_Type3 == 'apt_building'] <- 'home'
dat$Building_Type3 <- factor(dat$Building_Type3)                  

# wesBT <- data.frame(cols=c('cornflowerblue', 'darkolivegreen', 'darkorchid', 'darkorange'), 
#                     levs=levels(factor(dat$Building_Type3)))
wesBT <- data.frame(cols=c(wes_palette('Darjeeling', 5)[c(1,3,5)], 'darkorchid')[c(1,4,2,3)], 
                    levs=levels(factor(dat$Building_Type3)))
wesMAT <- data.frame(cols=c(wes_palette('Darjeeling', 5)[c(1,2,3)], 'darkorchid'), 
                     levs=levels(factor(dat$Matrix)))
table(dat$Matrix)
dat$colBuildingType <- as.character(wesBT$cols[match(dat$Building_Type, wesBT$levs)])
dat$colMatrix <- as.character(wesMAT$cols[match(dat$Matrix, wesMAT$levs)])




```


names(dat)
table(factor(dat$Matrix))
table(factor(dat$Building_Type))
table(factor(dat$Building_Type2))

Royal1, 1
Cavalcanti, 5
Rushmore, 5





```{r whiteBoxFunction}
# white lines with segments, fat lwd. 

whiteLines <- function(coords=NULL, poss=NULL, Names=NULL, alpha=.8, start=0.02, end=0.013, lwd=14) {
  llen <- nchar(Names)*end
  for(i in 1:length(poss)) {
    if (poss[i] == 4) {
      segments(x0=coords[i, 1] + start, 
               y0=coords[i, 2], 
               x1=coords[i, 1] + llen[i], 
               y1=coords[i, 2], 
               col=alpha('white', alpha=alpha), lwd=14)
    }
    if (poss[i] == 2) {
      segments(x0=coords[i, 1] - start, 
               y0=coords[i, 2], 
               x1=coords[i, 1] - llen[i], 
               y1=coords[i, 2], 
               col=alpha('white', alpha=alpha), lwd=14)
    }
  }
}    

# whiteLines(coords=ordUNI[, c(1,2)], poss=posUNI)
```




```{r plotOne, eval=FALSE}
pdf('ordUni.pdf', useDingbats=FALSE, height=6, width=6)
par(mar=c(2, 2, 1, 1))

plot(ordUNI[, c(1, 2)], pch=16, col=alpha(dat$col, alpha=.9), cex=1, # col='gray20', 
     bty='l', ann=FALSE, xaxt='n', yaxt='n', las=1)
mtext('PC1 (7.3%)', side=1, adj=1, line=.1)
mtext('PC2 (6.1%)', side=2, adj=1, line=.2)
# rect(-1, -1, 1, 1, col='gray')
whiteLines(coords=taxaUNI[, c(1,2)], poss=posUNI, Names=namesUNI, alpha=.8)
points(taxaUNI[, c(1, 2)], pch=24, bg='red', col='black', cex=1.5, lwd=2)
text(taxaUNI[, c(1,2)], labels=namesUNI, pos=posUNI, cex=.7, font=4)

dev.off()
```




```{r plot3ordinations}
pdf('ordThree.pdf', useDingbats=FALSE, height=10, width=10)
par(mfrow=c(2, 2), mar=c(2, 2, 1, 1))

# plot(ordBC[, c(1, 2)], pch=21, bg=alpha(dat$col, alpha=.7), 
#      col=alpha(dat$col, alpha=.9), cex=1.2, # col='gray20', 
#      bty='l', ann=FALSE, xaxt='n', yaxt='n')
# mtext('PC1 (7.4%)', side=1, adj=1, line=.1)
# mtext('PC2 (5.2%)', side=2, adj=1, line=.2)
# par(las=0, xpd=TRUE)
# mtext('   (a)', adj=0, line=-1, font=2, col='gray20', cex=1.4)
# # rect(-1, -1, 1, 1, col='gray')
# whiteLines(coords=taxaBC[, c(1,2)], poss=posBC, Names=namesBC, alpha=.8)
# points(taxaBC[, c(1, 2)], pch=24, bg='red', col='black', cex=1.5, lwd=2)
# text(taxaBC[, c(1,2)], labels=namesBC, pos=posBC, cex=.7, font=4)

plot(ordUNI[, c(1, 2)], pch=21, bg=alpha(dat$col, alpha=.7), 
     col=alpha(dat$col, alpha=.9), cex=1.2, # col='gray20', 
     bty='l', ann=FALSE, xaxt='n', yaxt='n', las=1)
mtext('PC1 (7.3%)', side=1, adj=1, line=.1)
mtext('PC2 (6.1%)', side=2, adj=1, line=.2)
mtext('   (a)', adj=0, line=-1, font=2, col='gray20', cex=1.4)
# rect(-1, -1, 1, 1, col='gray')
whiteLines(coords=taxaUNI[, c(1,2)], poss=posUNI, Names=namesUNI, alpha=.8)
points(taxaUNI[, c(1, 2)], pch=24, bg='red', col='black', cex=1.5, lwd=2)
text(taxaUNI[, c(1,2)], labels=namesUNI, pos=posUNI, cex=.7, font=4)

# exclude sources
studyNamesDFBE <- studyNamesDF[-c(2, 5, 6, 7, 11, 12, 23), ]

plot(0,0, axes=FALSE, ann=FALSE, type='n', xaxs='i', yaxs='i')
legend('bottomleft', legend=studyNamesDFBE$refNamesDisplay, pch=16, 
       pt.cex=3, cex=1.5, col=as.character(studyNamesDFBE$refCol), 
       bg='gray96', box.col='gray90')

purple <- which(dat$colMatrix == 'darkorchid')
plot(ordUNI[, c(1, 2)], pch=21, bg=alpha(dat$colMatrix, alpha=.7), 
     col=alpha(dat$colMatrix, alpha=.9), cex=1.2, # col='gray20', 
     bty='l', ann=FALSE, xaxt='n', yaxt='n', las=1)
points(ordUNI[, c(1, 2)][purple, ], 
       pch=21, bg=alpha(dat$colMatrix[purple], alpha=.7), 
       col=alpha(dat$colMatrix[purple], alpha=.9), cex=1.2)
mtext('PC1 (7.3%)', side=1, adj=1, line=.1)
mtext('PC2 (6.1%)', side=2, adj=1, line=.2)
mtext('   (b)', adj=0, line=-1, font=2, col='gray20', cex=1.4)
legend('bottomright', legend=wesMAT$levs, col=as.character(wesMAT$cols), 
       pch=16, pt.cex=2.5, box.col='gray90', bg='gray96', cex=1.3)
par(xpd=FALSE)
abline(h=par()$usr[3])

plot(ordUNI[, c(1, 2)], pch=21, bg=alpha(dat$colBuildingType, alpha=.7), 
     col=alpha(dat$colBuildingType, alpha=.9), cex=1.2, # col='gray20', 
     bty='l', ann=FALSE, xaxt='n', yaxt='n', las=1)
mtext('PC1 (7.3%)', side=1, adj=1, line=.1)
mtext('PC2 (6.1%)', side=2, adj=1, line=.2)
mtext('   (c)', adj=0, line=-1, font=2, col='gray20', cex=1.4)
legend('bottomright', legend=wesBT$levs, col=as.character(wesBT$cols), 
       pch=16, pt.cex=2.5, box.col='gray90', bg='gray96', cex=1.3)
abline(h=par()$usr[3])





dev.off()


```



```{r plotBC, eval=FALSE}
pdf('ordBC.pdf', useDingbats=FALSE, height=6, width=10)
par(mfrow=c(1, 2), mar=c(2, 2, 1, 1))

plot(ordBC[, c(1, 2)], pch=16, col=alpha(dat$col, alpha=.9), cex=1, # col='gray20', 
     bty='l', ann=FALSE, xaxt='n', yaxt='n')
mtext('PC1 (7.4%)', side=1, adj=1, line=.1)
mtext('PC2 (5.2%)', side=2, adj=1, line=.2)
par(las=0, xpd=TRUE)
mtext('   (a)', adj=0, line=-1, font=2, col='gray20', cex=1.4)
# rect(-1, -1, 1, 1, col='gray')
whiteLines(coords=taxaBC[, c(1,2)], poss=posBC, Names=namesBC, alpha=.8)
points(taxaBC[, c(1, 2)], pch=24, bg='red', col='black', cex=1.5, lwd=2)
text(taxaBC[, c(1,2)], labels=namesBC, pos=posBC, cex=.7, font=4)

plot(0,0, axes=FALSE, ann=FALSE, type='n', xaxs='i', yaxs='i')
legend('bottomleft', legend=studyNamesDFBE$refNamesDisplay, pch=16, 
       pt.cex=3, cex=1.5, col=as.character(studyNamesDFBE$refCol), 
       bg='gray96', box.col='gray90')


dev.off()
```



